---
title: "Manuscript10_6"
author: "Ashley Banuelos"
date: "2025-10-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages
```{r}
#Packages 
library(lme4)
library(emmeans)
library(tidyverse)
library(rstatix)
library(ggplot2)
library(performance)
library(ggpattern)
library(rlang)
library(flextable)
library(writexl)
```

Files
```{r}
#Files for analysis

#Climate variability across California 
climate <- read.csv("/Users/ashban/OneDrive - Colostate/Documents/CSU/Paustian/Point Blue/Manuscripts/Publishing_drafts_data/SOC_practices/SOC_R/alldata312.csv")
head(climate)

#Restored vs Unrestored 
socf <- read.csv("/Users/ashban/OneDrive - Colostate/Documents/CSU/Paustian/Point Blue/Lab work/updatedsheets/working/alldataWSL.csv")
socfnona <- socf[!is.na(socf$maomperc), ]
socfnona <- socfnona[!is.na(socfnona$corrcpom), ]
socfnona <- socfnona[!is.na(socfnona$bulkperc1), ]
socfnona<- socfnona[!is.na(socfnona$per_clay), ]
head(socfnona)
soc_riparian <- socfnona%>%
  filter(PlantingType.y == "Riparian")

soc_range <- socfnona%>%
  filter(PlantingType.y == "RangePlanting")

soc_tree <- socfnona%>%
  filter(PlantingType.y == "UplandOak")



# comparison 
soc_com <- read.csv("/Users/ashban/OneDrive - Colostate/Documents/CSU/Paustian/Point Blue/Lab work/updatedsheets/working/alldataWSL.csv")
soc_com_n<- soc_com[!is.na(soc_com$maoc_stock), ]
soc_com_n <- soc_com_n[!is.na(soc_com_n$poc_stock), ]
soc_com_n<- soc_com_n[!is.na(soc_com_n$per_clay), ]
soc_com_n <- soc_com_n%>%
  filter(trt == "T")

#canopy
fcanopy <- read.csv("/Users/ashban/OneDrive - Colostate/Documents/CSU/Paustian/Point Blue/Manuscripts/Publishing_drafts_data/SOC_practices/SOC_R/alldata312.csv")
fcanopy$Canopy<-ifelse(fcanopy$CanopyCover_0to4 > 0, "yes", "no")
fcanopy$Canopy<-as.factor(fcanopy$Canopy)
levels(fcanopy$Canopy)[levels(fcanopy$Canopy)=="yes"] <- "Under Canopy"
levels(fcanopy$Canopy)[levels(fcanopy$Canopy)=="no"] <- "Outside Canopy"
fcanopynona<- fcanopy[!is.na(fcanopy$Canopy), ]

```

Climate Differences
```{R}
#Climate differences 

# Step 1: Average pseudo-replicates within each Ranch pair first
pair_avg <- climate %>%
  group_by(pair, climate) %>%   # replace climate_group with your actual grouping column name
  summarise(
    soc_Mg.ha  = mean(soc_Mg.ha,  na.rm = TRUE),
    poc_stock  = mean(poc_stock,  na.rm = TRUE),
    maoc_stock = mean(maoc_stock, na.rm = TRUE),
    .groups = "drop"
  )

# Step 2: Run ANOVAs on pair averages with log transformation
aov_soc  <- aov(log(soc_Mg.ha)  ~ climate, data = pair_avg)
aov_poc  <- aov(log(poc_stock)  ~ climate, data = pair_avg)
aov_maoc <- aov(log(maoc_stock) ~ climate, data = pair_avg)

# Step 3: Check assumptions
check_model(aov_soc,  check = c("normality", "qq"))
check_model(aov_poc,  check = c("normality", "qq"))
check_model(aov_maoc, check = c("normality", "qq"))

# Step 4: Results
summary(aov_soc)
summary(aov_poc)
summary(aov_maoc)




```

TRT BY CLIMATE AND GROUP TRT RESULTS: No significant difference 

```{r}
#PER SOC 

lm_p <- lmer(log(bulkperc1)~ climate*trt+trt*age.group+per_clay+(1|pair)+(1|ranch.pract), data= socfnona)
emmeans(lm_p, revpairwise~ trt|climate)
emmeans(lm_p, revpairwise~ trt|age.group)

#Per POC 
lm_p <- lmer(log(corrcpom)~ climate*trt+trt*age.group+per_clayper_clay+(1|pair)+(1|ranch.pract), data= socfnona)
emmeans(lm_p, revpairwise~ trt|climate)
emmeans(lm_p, revpairwise~ trt|age.group)

#CHECK FOR ASSUMPTIONS 
check_model(lm_p, check=c("linearity", "homogeneity"))
summary(lm_p)

#PER MAOC
#Per POC 
lm_p <- lmer(log(maomperc)~ climate*trt+trt*age.group+per_clayper_clay+(1|pair)+(1|ranch.pract), data= socfnona)
emmeans(lm_p, revpairwise~ trt|climate)
emmeans(lm_p, revpairwise~ trt|age.group)


#CHECK FOR ASSUMPTIONS 
check_model(lm_m, check=c("linearity", "homogeneity"))
summary(lm_m)


```
Export climate results 
```{r}
#### export results data tables


#### export results data tables

# bulk C - by climate and age group
bulk_climate_tbl <- as.data.frame(emmeans(lm_bulk, revpairwise ~ trt | climate)$contrasts)
bulk_age_tbl     <- as.data.frame(emmeans(lm_bulk, revpairwise ~ trt | age.group)$contrasts)

# corrcpom - by climate and age group
corr_climate_tbl <- as.data.frame(emmeans(lm_corr, revpairwise ~ trt | climate)$contrasts)
corr_age_tbl     <- as.data.frame(emmeans(lm_corr, revpairwise ~ trt | age.group)$contrasts)

# maomperc - by climate and age group
maom_climate_tbl <- as.data.frame(emmeans(lm_maom, revpairwise ~ trt | climate)$contrasts)
maom_age_tbl     <- as.data.frame(emmeans(lm_maom, revpairwise ~ trt | age.group)$contrasts)


doc <- read_docx() |>
  body_add_par("Treatment × Climate and Age Group Effects on Soil Carbon Pools",
               style = "heading 1") |>

  body_add_par("Bulk C (%) — by Climate", style = "heading 2") |>
  body_add_flextable(flextable(bulk_climate_tbl)) |>
  body_add_par("Bulk C (%) — by Age Group", style = "heading 2") |>
  body_add_flextable(flextable(bulk_age_tbl)) |>

  body_add_par("Corrected POM (corrcpom) — by Climate", style = "heading 2") |>
  body_add_flextable(flextable(corr_climate_tbl)) |>
  body_add_par("Corrected POM (corrcpom) — by Age Group", style = "heading 2") |>
  body_add_flextable(flextable(corr_age_tbl)) |>

  body_add_par("MAOM (maomperc) — by Climate", style = "heading 2") |>
  body_add_flextable(flextable(maom_climate_tbl)) |>
  body_add_par("MAOM (maomperc) — by Age Group", style = "heading 2") |>
  body_add_flextable(flextable(maom_age_tbl))

print(doc, target = "TrtContrasts_Climate_AgeGroup.docx")
  

```

RIPARIAN SOC, POC, MAOC 

```{r}

#Carbon over time by trt (SOC)
lm_p <- lmer(log(bulkperc1)~age_2023*trt+per_clay+(1|pair/ranch.pract/RanchName.x), data= soc_riparian)
emtrends(lm_p, ~ trt, var = "age_2023")
emtrends(lm_p, pairwise ~ trt, var = "age_2023")
#PER SOC 

lm_p <- lmer(log(bulkperc1)~ climate*trt+age.group*trt+per_clay+(1|pair/ranch.pract/RanchName.x), data= soc_riparian)
emmeans(lm_p, revpairwise~ trt|climate)
emmeans(lm_p, revpairwise~ trt|age.group)


emmeans(lm_p, revpairwise~ trt|climate)
emmeans(lm_p, revpairwise~ trt|age.group)

#Per POC 
lm_p <- lmer(log(corrcpom)~ climate*trt+age.group*trt+per_clay+(1|pair/ranch.pract/RanchName.x), data= soc_riparian)
emmeans(lm_p, revpairwise~ trt|climate)
emmeans(lm_p, revpairwise~ trt|age.group)

#CHECK FOR ASSUMPTIONS 
check_model(lm_p, check=c("linearity", "homogeneity"))
summary(lm_p)

#PER MAOC
lm_m <- lmer(log(maomperc)~ climate*trt+age.group*trt+per_clay+(1|pair/ranch.pract/RanchName.x), data= soc_riparian)
emmeans(lm_m, revpairwise~ trt|climate)
emmeans(lm_m, revpairwise~ trt|age.group)

#CHECK FOR ASSUMPTIONS 
check_model(lm_m, check=c("linearity", "homogeneity"))
summary(lm_m)



```
Export riparian results  
```{r}
#### export results data tables - riparian

# bulk C - by climate and age group
bulk_climate_tbl <- as.data.frame(emmeans(lm_p, revpairwise ~ trt | climate)$contrasts)
bulk_age_tbl     <- as.data.frame(emmeans(lm_p, revpairwise ~ trt | age.group)$contrasts)

# corrcpom - by climate and age group
corr_climate_tbl <- as.data.frame(emmeans(lm_p, revpairwise ~ trt | climate)$contrasts)
corr_age_tbl     <- as.data.frame(emmeans(lm_p, revpairwise ~ trt | age.group)$contrasts)

# maomperc - by climate and age group
maom_climate_tbl <- as.data.frame(emmeans(lm_m, revpairwise ~ trt | climate)$contrasts)
maom_age_tbl     <- as.data.frame(emmeans(lm_m, revpairwise ~ trt | age.group)$contrasts)

library(flextable)
library(officer)

doc <- read_docx() |>
  body_add_par("Treatment × Climate and Age Group Effects on Soil Carbon Pools — Riparian",
               style = "heading 1") |>

  body_add_par("Bulk C (%) — by Climate", style = "heading 2") |>
  body_add_flextable(flextable(bulk_climate_tbl)) |>
  body_add_par("Bulk C (%) — by Age Group", style = "heading 2") |>
  body_add_flextable(flextable(bulk_age_tbl)) |>

  body_add_par("Corrected POM (corrcpom) — by Climate", style = "heading 2") |>
  body_add_flextable(flextable(corr_climate_tbl)) |>
  body_add_par("Corrected POM (corrcpom) — by Age Group", style = "heading 2") |>
  body_add_flextable(flextable(corr_age_tbl)) |>

  body_add_par("MAOM (maomperc) — by Climate", style = "heading 2") |>
  body_add_flextable(flextable(maom_climate_tbl)) |>
  body_add_par("MAOM (maomperc) — by Age Group", style = "heading 2") |>
  body_add_flextable(flextable(maom_age_tbl))

print(doc, target = "TrtContrasts_Riparian_Climate_AgeGroup.docx")
```

Upland Oak SOC, POC, MAOC 

```{r}
#carbon over time by trt (SOC)
lm_p <- lmer(log(bulkperc1)~age_2023*trt+per_clay+(1|pair/ranch.pract/RanchName.x), data= soc_tree)
emtrends(lm_p, ~ trt, var = "age_2023")
emtrends(lm_p, pairwise ~ trt, var = "age_2023")

#PER SOC 


lm_p <- lmer(log(bulkperc1)~ climate*trt+age.group*trt+per_clay+(1|ranch.pract/pair), data= soc_tree)
emmeans(lm_p, revpairwise~ trt|climate)
emmeans(lm_p, revpairwise~ trt|age.group)


#Per POC 
lm_p <- lmer(log(corrcpom)~ climate*trt+age.group*trt+per_clay+(1|ranch.pract/pair), data= soc_tree)
emmeans(lm_p, revpairwise~ trt|climate)
emmeans(lm_p, revpairwise~ trt|age.group)

#CHECK FOR ASSUMPTIONS 
check_model(lm_p, check=c("linearity", "homogeneity"))
summary(lm_p)

#PER MAOC
lm_m <- lmer(log(maomperc)~ climate*trt+age.group*trtper_clay+(1|ranch.pract/pair), data= soc_tree)
emmeans(lm_m, revpairwise~ trt|climate)
emmeans(lm_m, revpairwise~ trt|age.group)

#CHECK FOR ASSUMPTIONS 
check_model(lm_m, check=c("linearity", "homogeneity"))
summary(lm_m)



```

Export oak results 
```{r}
#### export results data tables - tree

# bulk C - by climate and age group
bulk_climate_tbl <- as.data.frame(emmeans(lm_p, revpairwise ~ trt | climate)$contrasts)
bulk_age_tbl     <- as.data.frame(emmeans(lm_p, revpairwise ~ trt | age.group)$contrasts)

# corrcpom - by climate and age group
corr_climate_tbl <- as.data.frame(emmeans(lm_p, revpairwise ~ trt | climate)$contrasts)
corr_age_tbl     <- as.data.frame(emmeans(lm_p, revpairwise ~ trt | age.group)$contrasts)

# maomperc - by climate and age group
maom_climate_tbl <- as.data.frame(emmeans(lm_m, revpairwise ~ trt | climate)$contrasts)
maom_age_tbl     <- as.data.frame(emmeans(lm_m, revpairwise ~ trt | age.group)$contrasts)

library(flextable)
library(officer)

doc <- read_docx() |>
  body_add_par("Treatment × Climate and Age Group Effects on Soil Carbon Pools — Tree",
               style = "heading 1") |>

  body_add_par("Bulk C (%) — by Climate", style = "heading 2") |>
  body_add_flextable(flextable(bulk_climate_tbl)) |>
  body_add_par("Bulk C (%) — by Age Group", style = "heading 2") |>
  body_add_flextable(flextable(bulk_age_tbl)) |>

  body_add_par("Corrected POM (corrcpom) — by Climate", style = "heading 2") |>
  body_add_flextable(flextable(corr_climate_tbl)) |>
  body_add_par("Corrected POM (corrcpom) — by Age Group", style = "heading 2") |>
  body_add_flextable(flextable(corr_age_tbl)) |>

  body_add_par("MAOM (maomperc) — by Climate", style = "heading 2") |>
  body_add_flextable(flextable(maom_climate_tbl)) |>
  body_add_par("MAOM (maomperc) — by Age Group", style = "heading 2") |>
  body_add_flextable(flextable(maom_age_tbl))

print(doc, target = "TrtContrasts_Tree_Climate_AgeGroup.docx")

```

Perennial Grass SOC, POC, MAOC 

```{r}

#Carbon over time by trt (SOC)

lm_p <- lmer(log(bulkperc1)~age_2023*trt+per_clay+(1|pair/ranch.pract/RanchName.x), data= soc_range)
emtrends(lm_p, ~ trt, var = "age_2023")
emtrends(lm_p, pairwise ~ trt, var = "age_2023")
#PER SOC 

lm_p <- lmer(log(bulkperc1)~ climate*trt+age.group*trt+per_clay+(1|ranch.pract/pair), data= soc_range)
emmeans(lm_p, revpairwise~ trt|climate)
emmeans(lm_p, revpairwise~ trt|age.group)


#Per POC 
lm_p <- lmer(log(corrcpom)~ climate*trt+age.group*trt+per_clay+(1|ranch.pract/pair), data= soc_range)
emmeans(lm_p, revpairwise~ trt|climate)
emmeans(lm_p, revpairwise~ trt|age.group)

#CHECK FOR ASSUMPTIONS 
check_model(lm_p, check=c("linearity", "homogeneity"))
summary(lm_p)

#PER MAOC
lm_m <- lmer(log(maomperc)~ climate*trt+age.group*trt+per_clay+(1|pair/ranch.pract), data= soc_range)
emmeans(lm_m, revpairwise~ trt|climate)
emmeans(lm_m, revpairwise~ trt|age.group)

#CHECK FOR ASSUMPTIONS 
check_model(lm_m, check=c("linearity", "homogeneity"))
summary(lm_m)

```

Export perennial results 
```{r}
#### export results data tables - range

# bulk C - by climate and age group
bulk_climate_tbl <- as.data.frame(emmeans(lm_p, revpairwise ~ trt | climate)$contrasts)
bulk_age_tbl     <- as.data.frame(emmeans(lm_p, revpairwise ~ trt | age.group)$contrasts)

# corrcpom - by climate and age group
corr_climate_tbl <- as.data.frame(emmeans(lm_p, revpairwise ~ trt | climate)$contrasts)
corr_age_tbl     <- as.data.frame(emmeans(lm_p, revpairwise ~ trt | age.group)$contrasts)

# maomperc - by climate and age group
maom_climate_tbl <- as.data.frame(emmeans(lm_m, revpairwise ~ trt | climate)$contrasts)
maom_age_tbl     <- as.data.frame(emmeans(lm_m, revpairwise ~ trt | age.group)$contrasts)

library(flextable)
library(officer)

doc <- read_docx() |>
  body_add_par("Treatment × Climate and Age Group Effects on Soil Carbon Pools — Range",
               style = "heading 1") |>

  body_add_par("Bulk C (%) — by Climate", style = "heading 2") |>
  body_add_flextable(flextable(bulk_climate_tbl)) |>
  body_add_par("Bulk C (%) — by Age Group", style = "heading 2") |>
  body_add_flextable(flextable(bulk_age_tbl)) |>

  body_add_par("Corrected POM (corrcpom) — by Climate", style = "heading 2") |>
  body_add_flextable(flextable(corr_climate_tbl)) |>
  body_add_par("Corrected POM (corrcpom) — by Age Group", style = "heading 2") |>
  body_add_flextable(flextable(corr_age_tbl)) |>

  body_add_par("MAOM (maomperc) — by Climate", style = "heading 2") |>
  body_add_flextable(flextable(maom_climate_tbl)) |>
  body_add_par("MAOM (maomperc) — by Age Group", style = "heading 2") |>
  body_add_flextable(flextable(maom_age_tbl))

print(doc, target = "TrtContrasts_Range_Climate_AgeGroup.docx")

```

COMPARISON RESULTS: 
SOC STOCK - No significant difference 

POC STOCK - 
  - AGE: Medium: Range planting higher compared tree (69%, 13.1, p=0.03)

  - CLIMATE: Dry - perennial has higher POC compared to riparian (44%, p = 0.05) and oaks (52%, p = 0.02).
  
  
MAOC STOCK -
  - CLIMATE: Wet - Oaks higher than range (0.65, 37,  P= 0.08)
  
Comparisons of conservation practices 
```{r}
#COMPARISONS 

#SOC stock 

soc_climate <- lmer(log(soc_Mg.ha)~ climate*PlantingType.y+per_clay+(1|pair), data= soc_com_n)
emmeans(soc_climate , revpairwise~ PlantingType.y|climate)

soc_age <- lmer(log(soc_Mg.ha)~age.group*PlantingType.y+per_clay+(1|pair), data= soc_com_n)
emmeans(soc_age, revpairwise~ PlantingType.y|age.group)


#POC stock 
poc_climate  <- lmer(log(poc_stock)~ climate*PlantingType.y+per_clay+(1|pair), data= soc_com_n)
emmeans(poc_climate , pairwise~ PlantingType.y|climate)
check_model(poc_climate, check=c("linearity", "homogeneity"))

poc_age  <- lmer(log(poc_stock)~age.group*PlantingType.y+per_clay+(1|pair), data= soc_com_n)
emmeans(poc_age, revpairwise~ PlantingType.y|age.group)
check_model(poc_age, check=c("linearity", "homogeneity"))

#MAOC STOCK 
maoc_climate <- lmer(log(maoc_stock)~ climate*PlantingType.y+per_clay+(1|pair), data= soc_com_n)
emmeans(maoc_climate, revpairwise~ PlantingType.y|climate)

maoc_age<- lmer(log(maoc_stock)~age.group*PlantingType.y+per_clay+(1|pair), data= soc_com_n)
emmeans(maoc_age, revpairwise~ PlantingType.y|age.group)

#CHECK FOR ASSUMPTIONS 
check_model(maoc_climate, check=c("linearity", "homogeneity"))
check_model(maoc_age, check=c("linearity", "homogeneity"))
summary(maoc_climate)

```

SOC over time WITH reference 
```{r}
#### Analysis - trends over time ####

# Set age_2023 to 0 where trt is "T"
socfnona_0 <- socfnona %>%
  mutate(age_2023 = ifelse(trt == "C", 0, age_2023))

# SOC
soc_time <- lm(log(soc_Mg.ha) ~ PlantingType.y * climate * age_2023 + per_clay,
               data = socfnona_0)
summary(soc_time)
emtrends(soc_time, ~ PlantingType.y | climate, var = "age_2023")

# MAOC
maoc_time <- lm(log(maoc_stock) ~ PlantingType.y * climate * age_2023 + per_clay,
                data = socfnona_0)
anova(maoc_time)
emtrends(maoc_time, ~ PlantingType.y | climate, var = "age_2023")

# POC
poc_time <- lm(log(poc_stock) ~ PlantingType.y * climate * age_2023 + per_clay,
               data = socfnona_0)
anova(poc_time)
emtrends(poc_time, ~ PlantingType.y | climate, var = "age_2023")

# Extract trends with CI
soc_ci <- emtrends(soc_time, ~ PlantingType.y | climate, var = "age_2023") %>%
  as.data.frame()

# Extract trends with p-values
soc_p <- emtrends(soc_time, ~ PlantingType.y | climate, var = "age_2023") %>%
  test() %>% as.data.frame()

# Join together
soc_trends <- left_join(soc_ci, soc_p, by = c("PlantingType.y", "climate", "age_2023.trend", "SE", "df")) %>%
  mutate(Fraction = "SOC")

# Repeat for MAOC
maoc_ci <- emtrends(maoc_time, ~ PlantingType.y | climate, var = "age_2023") %>% as.data.frame()
maoc_p  <- emtrends(maoc_time, ~ PlantingType.y | climate, var = "age_2023") %>% test() %>% as.data.frame()
maoc_trends <- left_join(maoc_ci, maoc_p, by = c("PlantingType.y", "climate", "age_2023.trend", "SE", "df")) %>%
  mutate(Fraction = "MAOC")

# Repeat for POC
poc_ci <- emtrends(poc_time, ~ PlantingType.y | climate, var = "age_2023") %>% as.data.frame()
poc_p  <- emtrends(poc_time, ~ PlantingType.y | climate, var = "age_2023") %>% test() %>% as.data.frame()
poc_trends <- left_join(poc_ci, poc_p, by = c("PlantingType.y", "climate", "age_2023.trend", "SE", "df")) %>%
  mutate(Fraction = "POC")

# Combine and clean
all_trends <- bind_rows(soc_trends, maoc_trends, poc_trends) %>%
  select(Fraction, climate, PlantingType.y, age_2023.trend, SE, lower.CL, upper.CL, t.ratio, p.value) %>%
  rename(
    Fraction = Fraction,
    Climate  = climate,
    Practice = PlantingType.y,
    Slope    = age_2023.trend,
    P.value  = p.value
  ) %>%
  mutate(
    Slope    = round(Slope,    4),
    SE       = round(SE,       4),
    lower.CL = round(lower.CL, 4),
    upper.CL = round(upper.CL, 4),
    P.value  = round(P.value,  4)
  )

write_xlsx(all_trends, "carbon_trends.xlsx")
```
Extracting data for carbon over time 
```{r}
soc_trends <- emtrends(soc_time, ~ PlantingType.y | climate, var = "age_2023") %>%
  test(side = "two-sided") %>%
  as.data.frame() %>%
  mutate(Fraction = "SOC")
get_trends <- function(model, fraction) {
  emtrends(model, ~ PlantingType.y | climate, var = "age_2023") %>%
    confint() %>%
    as.data.frame() %>%
    left_join(
      emtrends(model, ~ PlantingType.y | climate, var = "age_2023") %>%
        test() %>%
        as.data.frame() %>%
        select(PlantingType.y, climate, t.ratio, p.value),
      by = c("PlantingType.y", "climate")
    ) %>%
    mutate(Fraction = fraction)
}

all_trends <- bind_rows(
  get_trends(soc_time,  "SOC"),
  get_trends(maoc_time, "MAOC"),
  get_trends(poc_time,  "POC")
) %>%
  select(Fraction, climate, PlantingType.y, age_2023.trend, SE, lower.CL, upper.CL, t.ratio, p.value) %>%
  rename(
    Climate  = climate,
    Practice = PlantingType.y,
    Slope    = age_2023.trend,
    P.value  = p.value
  ) %>%
  mutate(across(where(is.numeric), ~ round(.x, 4)))

write_xlsx(all_trends, "carbon_trends.xlsx")
```

Extract R² and model fit statistics
```{r} 
library(broom)
library(dplyr)
library(writexl)

####extract R2
model_fit <- data.frame(
  response     = c("SOC", "MAOC", "POC"),
  r_squared    = c(summary(soc_time)$r.squared,
                   summary(maoc_time)$r.squared,
                   summary(poc_time)$r.squared),
  adj_r_squared = c(summary(soc_time)$adj.r.squared,
                    summary(maoc_time)$adj.r.squared,
                    summary(poc_time)$adj.r.squared),
  F_statistic  = c(summary(soc_time)$fstatistic[1],
                   summary(maoc_time)$fstatistic[1],
                   summary(poc_time)$fstatistic[1]),
  df1          = c(summary(soc_time)$fstatistic[2],
                   summary(maoc_time)$fstatistic[2],
                   summary(poc_time)$fstatistic[2]),
  df2          = c(summary(soc_time)$fstatistic[3],
                   summary(maoc_time)$fstatistic[3],
                   summary(poc_time)$fstatistic[3])
) %>%
   mutate(
    p_raw     = pf(F_statistic, df1, df2, lower.tail = FALSE),
    # ✅ Format p-value as text so it never rounds to 0
    p_value   = ifelse(p_raw < 0.001, "< 0.001", round(p_raw, 4)),
    r_squared     = round(r_squared, 4),
    adj_r_squared = round(adj_r_squared, 4),
    F_statistic   = round(F_statistic, 4)
  ) %>%
  select(-p_raw)
  

print(model_fit)

# ---- Add to Excel export ----
write_xlsx(
  list(
    "Model_fit"    = model_fit,
    "ANOVA_all"    = anova_all,
    "ANOVA_SOC"    = anova_soc,
    "ANOVA_MAOC"   = anova_maoc,
    "ANOVA_POC"    = anova_poc,
    "Trends_all"   = emtrend_all,
    "Trends_SOC"   = emtrend_soc,
    "Trends_MAOC"  = emtrend_maoc,
    "Trends_POC"   = emtrend_poc
  ),
  path = "SOC_model_results.xlsx"
)

```


CANOPY - OVERALL 

RESULTS: 
 OVERALL SOC - 9% p=0.02
 OVERALL MOAC - 10% P=0.02
 OVERALL POC - 9% p=0.02

```{r}
###### without any interactions --  over all 
canopy_corr <-lmer(log(corrcpom)~ Canopy + per_clay  + (1|RanchName.x), data = fcanopynona)
emmeans(canopy_corr, revpairwise~ Canopy)


canopy_maoc <-lmer(log(maomperc)~ Canopy + per_clay  + (1|RanchName.x), data = fcanopynona)
emmeans(canopy_maoc, revpairwise~ Canopy)

canopy_soc <-lmer(log(bulkperc1 )~ Canopy + per_clay  + (1|RanchName.x), data = fcanopynona)
emmeans(canopy_soc, revpairwise~ Canopy)

```

Canopy by climate
```{r}
##### Canopy by climate 
library(lme4)
library(emmeans)

canopy_corr_clim <- lmer(
  log(corrcpom) ~ Canopy * climate + per_clay + (1 | RanchName.x),
  data = fcanopynona
)

canopy_maoc_clim <- lmer(
  log(maomperc) ~ Canopy * climate + per_clay + (1 | RanchName.x),
  data = fcanopynona
)

canopy_soc_clim <- lmer(
  log(bulkperc1) ~ Canopy * climate + per_clay + (1 | RanchName.x),
  data = fcanopynona
)



```

Export Canopy results
```{r}
###export results ### overall export  canopy 
library(officer)
library(flextable)

emm_corr <- emmeans(canopy_corr, revpairwise ~ Canopy)
emm_maoc <- emmeans(canopy_maoc, revpairwise ~ Canopy)
emm_soc  <- emmeans(canopy_soc,  revpairwise ~ Canopy)

corr_tbl <- as.data.frame(emm_corr$contrasts)
maoc_tbl <- as.data.frame(emm_maoc$contrasts)
soc_tbl  <- as.data.frame(emm_soc$contrasts)



canopy_results <- read_docx() |>
  body_add_par("Canopy Effects on Carbon Pools", style = "heading 1") |>

  body_add_par("Corrected POM", style = "heading 2") |>
  body_add_flextable(flextable(corr_tbl)) |>

  body_add_par("MAOM", style = "heading 2") |>
  body_add_flextable(flextable(maoc_tbl)) |>

  body_add_par("SOC", style = "heading 2") |>
  body_add_flextable(flextable(soc_tbl))

print(canopy_results, target = "Canopy_overall_results.docx")


```

CANOPY - CLIMATE

RESULTS:
 CLIMATE POC - DRY 11%   P = 0.03
 CLIMATE MAOC - WET 23% P=0.0002
 CLIMATE SOC - wet 19% p= 0.0001
 
```{R}
library(lme4)
library(emmeans)

canopy_corr_clim <- lmer(
  log(corrcpom) ~ Canopy * climate + per_clay + (1 | RanchName.x),
  data = fcanopynona
)

canopy_maoc_clim <- lmer(
  log(maomperc) ~ Canopy * climate + per_clay + (1 | RanchName.x),
  data = fcanopynona
)

canopy_soc_clim <- lmer(
  log(bulkperc1) ~ Canopy * climate + per_clay + (1 | RanchName.x),
  data = fcanopynona
)

emm_corr <- emmeans(canopy_corr_clim, revpairwise ~ Canopy | climate)
emm_maoc <- emmeans(canopy_maoc_clim, revpairwise ~ Canopy | climate)
emm_soc  <- emmeans(canopy_soc_clim,  revpairwise ~ Canopy | climate)


#CHECK FOR ASSUMPTIONS
check_model(canopy1, check=c("linearity", "homogeneity"))
```

export result canopy climate 
```{r}
####  export results data tables 

corr_tbl <- as.data.frame(emm_corr$contrasts)
maoc_tbl <- as.data.frame(emm_maoc$contrasts)
soc_tbl  <- as.data.frame(emm_soc$contrasts)

library(flextable)
library(officer)

doc <- read_docx() |>
  body_add_par("Canopy × Climate Effects on Soil Carbon Pools",
               style = "heading 1") |>

  body_add_par("Corrected POM (corrcpom)", style = "heading 2") |>
  body_add_flextable(flextable(corr_tbl)) |>

  body_add_par("MAOM (maomperc)", style = "heading 2") |>
  body_add_flextable(flextable(maoc_tbl)) |>

  body_add_par("SOC (%)", style = "heading 2") |>
  body_add_flextable(flextable(soc_tbl))

print(doc, target = "Canopy_by_Climate_emmeans.docx")



```

Cohen's d
```{r}

#### cohen's d ########################
# 0.03 
# Subset the data
x1 <- alldataWSL$bulkperc1[alldataWSL$trt == "T"]
x2 <- alldataWSL$bulkperc1[alldataWSL$trt == "C"]

# Means
mean1 <- mean(x1, na.rm = TRUE)
mean2 <- mean(x2, na.rm = TRUE)

# Standard deviations
sd1 <- sd(x1, na.rm = TRUE)
sd2 <- sd(x2, na.rm = TRUE)

# Sample sizes
n1 <- length(x1)
n2 <- length(x2)

# Pooled standard deviation
sp <- sqrt(((n1 - 1) * sd1^2 + (n2 - 1) * sd2^2) / (n1 + n2 - 2))

# Cohen's d
cohens_d <- (mean1 - mean2) / sp

# Print result
cat("Cohen's d:", cohens_d, "\n")


##### with confience interval 
# install.packages("effsize") # if not already installed
library(effsize)

# Cohen's d with confidence interval
d_result <- cohen.d(x1, x2, pooled = TRUE, hedges.correction = TRUE, conf.level = 0.95)
d_result




```

Export comparison results
```{r}

library(broom.mixed)
library(writexl)

# Function to extract fixed effects from each model
extract_model <- function(model, response, predictor_type) {
  tidy(model, effects = "fixed") %>%
    mutate(
      Response = response,
      Model    = predictor_type
    ) %>%
    select(Response, Model, term, estimate, std.error, statistic, p.value)
}

# Extract from each model
soc_clim_df  <- extract_model(soc_climate,  "SOC stock",  "Climate")
soc_age_df   <- extract_model(soc_age,      "SOC stock",  "Age Group")
poc_clim_df  <- extract_model(poc_climate,  "POC stock",  "Climate")
poc_age_df   <- extract_model(poc_age,      "POC stock",  "Age Group")
maoc_clim_df <- extract_model(maoc_climate, "MAOC stock", "Climate")
maoc_age_df  <- extract_model(maoc_age,     "MAOC stock", "Age Group")

# Combine all
all_models <- bind_rows(
  soc_clim_df, soc_age_df,
  poc_clim_df, poc_age_df,
  maoc_clim_df, maoc_age_df
) %>%
  rename(
    Term      = term,
    Estimate  = estimate,
    SE        = std.error,
    Statistic = statistic,
    P.value   = p.value
  ) %>%
  mutate(
    Estimate  = round(Estimate,  3),
    SE        = round(SE,        3),
    Statistic = round(Statistic, 3),
    P.value   = round(P.value,   4)
  )

# Export to Excel
write_xlsx(all_models, "model_results.xlsx")

```

Table by climate 
```{r}

## export table by climate (dry/mesic)

soc_table_fmt <- soc_com %>%
  group_by(climate) %>%
  summarise(
    n = n(),
        bulkperc1 = sprintf("%.2f ± %.2f",
                        mean(bulkperc1, na.rm = TRUE),
                        sd(bulkperc1, na.rm = TRUE) / sqrt(n)),

    corrcpom = sprintf("%.2f ± %.2f",
                       mean(corrcpom, na.rm = TRUE),
                       sd(corrcpom, na.rm = TRUE) / sqrt(n)),

    maomperc = sprintf("%.2f ± %.2f",
                       mean(maomperc, na.rm = TRUE),
                       sd(maomperc, na.rm = TRUE) / sqrt(n)),
    soc_Mg.ha = sprintf("%.2f ± %.2f",
                        mean(soc_Mg.ha, na.rm = TRUE),
                        sd(soc_Mg.ha, na.rm = TRUE) / sqrt(n)),
    
    poc_stock = sprintf("%.2f ± %.2f",
                        mean(poc_stock, na.rm = TRUE),
                        sd(poc_stock, na.rm = TRUE) / sqrt(n)),
    
    maoc_stock = sprintf("%.2f ± %.2f",
                         mean(maoc_stock, na.rm = TRUE),
                         sd(maoc_stock, na.rm = TRUE) / sqrt(n)),
    
    fmaoc = sprintf("%.2f ± %.2f",
                    mean(fmaoc, na.rm = TRUE),
                    sd(fmaoc, na.rm = TRUE) / sqrt(n)),
    
    per_sand = sprintf("%.1f ± %.1f",
                       mean(per_sand, na.rm = TRUE),
                       sd(per_sand, na.rm = TRUE) / sqrt(n)),
    
    per_clay = sprintf("%.1f ± %.1f",
                       mean(per_clay, na.rm = TRUE),
                       sd(per_clay, na.rm = TRUE) / sqrt(n)),
    
    per_silt = sprintf("%.1f ± %.1f",
                       mean(per_silt, na.rm = TRUE),
                       sd(per_silt, na.rm = TRUE) / sqrt(n)),
    
    pH = sprintf("%.2f ± %.2f",
                 mean(pH, na.rm = TRUE),
                 sd(pH, na.rm = TRUE) / sqrt(n)),
    
    BD_g.cm3 = sprintf("%.2f ± %.2f",
                       mean(BD_g.cm3, na.rm = TRUE),
                       sd(BD_g.cm3, na.rm = TRUE) / sqrt(n)),
    
  biomass_gm2 = sprintf("%.2f ± %.2f",
                       mean(biomass_gm2, na.rm = TRUE),
                       sd(biomass_gm2, na.rm = TRUE) / sqrt(n)),
    
    .groups = "drop"
  )

soc_table_fmt_n <- soc_table_fmt %>%
  group_by(climate) %>%
  mutate(n = n()) %>%
  ungroup()


soc_flex <- flextable(soc_table_fmt) %>%
  set_header_labels(
     climate = "Climate", n = "n",
    bulkperc1 = "Bulk C (%)",
    corrcpom = "cPOM (%)",
    maomperc = "MAOM (%)",
    soc_Mg.ha = "SOC (Mg ha⁻¹)",
    poc_stock = "POC stock (Mg ha⁻¹)",
    maoc_stock = "MAOC stock (Mg ha⁻¹)",
    fmaoc = "fMAOC",
    per_sand = "Sand (%)",
    per_clay = "Clay (%)",
    per_silt = "Silt (%)",
    pH = "pH",
    BD_g.cm3 = "Bulk density (g cm⁻³)",
     biomass_gm2 = "Herbaceous biomass (g m⁻²)"
  ) %>%
  align(align = "center", part = "all") %>%
  autofit()

save_as_docx(soc_flex, path = "clime_only.docx")
```

TRT
```{r}
#### Aridity level 
## export table by climate (dry/mesic)
soc_table_fmt <- soc_com %>%
  group_by(trt) %>%
  summarise(
    n = n(),
        bulkperc1 = sprintf("%.2f ± %.2f",
                        mean(bulkperc1, na.rm = TRUE),
                        sd(bulkperc1, na.rm = TRUE) / sqrt(n)),

    corrcpom = sprintf("%.2f ± %.2f",
                       mean(corrcpom, na.rm = TRUE),
                       sd(corrcpom, na.rm = TRUE) / sqrt(n)),

    maomperc = sprintf("%.2f ± %.2f",
                       mean(maomperc, na.rm = TRUE),
                       sd(maomperc, na.rm = TRUE) / sqrt(n)),
    soc_Mg.ha = sprintf("%.2f ± %.2f",
                        mean(soc_Mg.ha, na.rm = TRUE),
                        sd(soc_Mg.ha, na.rm = TRUE) / sqrt(n)),
    
    poc_stock = sprintf("%.2f ± %.2f",
                        mean(poc_stock, na.rm = TRUE),
                        sd(poc_stock, na.rm = TRUE) / sqrt(n)),
    
    maoc_stock = sprintf("%.2f ± %.2f",
                         mean(maoc_stock, na.rm = TRUE),
                         sd(maoc_stock, na.rm = TRUE) / sqrt(n)),
    
    fmaoc = sprintf("%.2f ± %.2f",
                    mean(fmaoc, na.rm = TRUE),
                    sd(fmaoc, na.rm = TRUE) / sqrt(n)),
    
    per_sand = sprintf("%.1f ± %.1f",
                       mean(per_sand, na.rm = TRUE),
                       sd(per_sand, na.rm = TRUE) / sqrt(n)),
    
    per_clay = sprintf("%.1f ± %.1f",
                       mean(per_clay, na.rm = TRUE),
                       sd(per_clay, na.rm = TRUE) / sqrt(n)),
    
    per_silt = sprintf("%.1f ± %.1f",
                       mean(per_silt, na.rm = TRUE),
                       sd(per_silt, na.rm = TRUE) / sqrt(n)),
    
    pH = sprintf("%.2f ± %.2f",
                 mean(pH, na.rm = TRUE),
                 sd(pH, na.rm = TRUE) / sqrt(n)),
    
    BD_g.cm3 = sprintf("%.2f ± %.2f",
                       mean(BD_g.cm3, na.rm = TRUE),
                       sd(BD_g.cm3, na.rm = TRUE) / sqrt(n)),
    
  biomass_gm2 = sprintf("%.2f ± %.2f",
                       mean(biomass_gm2, na.rm = TRUE),
                       sd(biomass_gm2, na.rm = TRUE) / sqrt(n)),
    
    .groups = "drop"
  )

soc_flex <- flextable(soc_table_fmt) %>%
  set_header_labels(
    trt = "Treatment",
    n = "n",
    bulkperc1 = "Bulk C (%)",
    corrcpom = "cPOM (%)",
    maomperc = "MAOM (%)",
    soc_Mg.ha = "SOC (Mg ha⁻¹)",
    poc_stock = "POC stock (Mg ha⁻¹)",
    maoc_stock = "MAOC stock (Mg ha⁻¹)",
    fmaoc = "fMAOC",
    per_sand = "Sand (%)",
    per_clay = "Clay (%)",
    per_silt = "Silt (%)",
    pH = "pH",
    BD_g.cm3 = "Bulk density (g cm⁻³)",
    biomass_gm2 = "Herbaceous biomass (g m⁻²)"
  ) %>%
  align(align = "center", part = "all") %>%
  flextable::autofit()

save_as_docx(soc_flex, path = "trt_only.docx")



```

Table practice climate and age group by trt
```{r}

### tables 
### practice climate and age group by trt

 library(flextable)
soc_table_fmt <- soc_com %>%
  group_by(PlantingType.y,  climate,age.group, trt) %>%
  summarise(
    n = n(),
        bulkperc1 = sprintf("%.2f ± %.2f",
                        mean(bulkperc1, na.rm = TRUE),
                        sd(bulkperc1, na.rm = TRUE) / sqrt(n)),

    corrcpom = sprintf("%.2f ± %.2f",
                       mean(corrcpom, na.rm = TRUE),
                       sd(corrcpom, na.rm = TRUE) / sqrt(n)),

    maomperc = sprintf("%.2f ± %.2f",
                       mean(maomperc, na.rm = TRUE),
                       sd(maomperc, na.rm = TRUE) / sqrt(n)),
    soc_Mg.ha = sprintf("%.2f ± %.2f",
                        mean(soc_Mg.ha, na.rm = TRUE),
                        sd(soc_Mg.ha, na.rm = TRUE) / sqrt(n)),
    
    poc_stock = sprintf("%.2f ± %.2f",
                        mean(poc_stock, na.rm = TRUE),
                        sd(poc_stock, na.rm = TRUE) / sqrt(n)),
    
    maoc_stock = sprintf("%.2f ± %.2f",
                         mean(maoc_stock, na.rm = TRUE),
                         sd(maoc_stock, na.rm = TRUE) / sqrt(n)),
    
    fmaoc = sprintf("%.2f ± %.2f",
                    mean(fmaoc, na.rm = TRUE),
                    sd(fmaoc, na.rm = TRUE) / sqrt(n)),
    
    per_sand = sprintf("%.1f ± %.1f",
                       mean(per_sand, na.rm = TRUE),
                       sd(per_sand, na.rm = TRUE) / sqrt(n)),
    
    per_clay = sprintf("%.1f ± %.1f",
                       mean(per_clay, na.rm = TRUE),
                       sd(per_clay, na.rm = TRUE) / sqrt(n)),
    
    per_silt = sprintf("%.1f ± %.1f",
                       mean(per_silt, na.rm = TRUE),
                       sd(per_silt, na.rm = TRUE) / sqrt(n)),
    
    pH = sprintf("%.2f ± %.2f",
                 mean(pH, na.rm = TRUE),
                 sd(pH, na.rm = TRUE) / sqrt(n)),
    
    BD_g.cm3 = sprintf("%.2f ± %.2f",
                       mean(BD_g.cm3, na.rm = TRUE),
                       sd(BD_g.cm3, na.rm = TRUE) / sqrt(n)),
    
  biomass_gm2 = sprintf("%.2f ± %.2f",
                       mean(biomass_gm2, na.rm = TRUE),
                       sd(biomass_gm2, na.rm = TRUE) / sqrt(n)),
    
    .groups = "drop"
  )

soc_flex <- flextable(soc_table_fmt) %>%
  set_header_labels(
    PlantingType.y = "Practice",
     climate = "Climate",
    age.group = "age group",
    trt = "trt",
    bulkperc1 = "Bulk C (%)",
    corrcpom = "cPOM (%)",
    maomperc = "MAOM (%)",
    soc_Mg.ha = "SOC (Mg ha⁻¹)",
    poc_stock = "POC stock (Mg ha⁻¹)",
    maoc_stock = "MAOC stock (Mg ha⁻¹)",
    fmaoc = "fMAOC",
    per_sand = "Sand (%)",
    per_clay = "Clay (%)",
    per_silt = "Silt (%)",
    pH = "pH",
    BD_g.cm3 = "Bulk density (g cm⁻³)",
     biomass_gm2 = "Herbaceous biomass (g m⁻²)"
  ) %>%
  align(align = "center", part = "all") %>%
  autofit()

save_as_docx(soc_flex, path = "cons_summary_clime_age.docx")



```

Table practice climate and trt only 
```{R}

### table practice climate and trt only  
library(flextable)
soc_table_fmt <- soc_com %>%
  group_by(PlantingType.y,  climate, trt) %>%
  summarise(
    n = n(),
        bulkperc1 = sprintf("%.2f ± %.2f",
                        mean(bulkperc1, na.rm = TRUE),
                        sd(bulkperc1, na.rm = TRUE) / sqrt(n)),

    corrcpom = sprintf("%.2f ± %.2f",
                       mean(corrcpom, na.rm = TRUE),
                       sd(corrcpom, na.rm = TRUE) / sqrt(n)),

    maomperc = sprintf("%.2f ± %.2f",
                       mean(maomperc, na.rm = TRUE),
                       sd(maomperc, na.rm = TRUE) / sqrt(n)),
    soc_Mg.ha = sprintf("%.2f ± %.2f",
                        mean(soc_Mg.ha, na.rm = TRUE),
                        sd(soc_Mg.ha, na.rm = TRUE) / sqrt(n)),
    
    poc_stock = sprintf("%.2f ± %.2f",
                        mean(poc_stock, na.rm = TRUE),
                        sd(poc_stock, na.rm = TRUE) / sqrt(n)),
    
    maoc_stock = sprintf("%.2f ± %.2f",
                         mean(maoc_stock, na.rm = TRUE),
                         sd(maoc_stock, na.rm = TRUE) / sqrt(n)),
    
    fmaoc = sprintf("%.2f ± %.2f",
                    mean(fmaoc, na.rm = TRUE),
                    sd(fmaoc, na.rm = TRUE) / sqrt(n)),
    
    per_sand = sprintf("%.1f ± %.1f",
                       mean(per_sand, na.rm = TRUE),
                       sd(per_sand, na.rm = TRUE) / sqrt(n)),
    
    per_clay = sprintf("%.1f ± %.1f",
                       mean(per_clay, na.rm = TRUE),
                       sd(per_clay, na.rm = TRUE) / sqrt(n)),
    
    per_silt = sprintf("%.1f ± %.1f",
                       mean(per_silt, na.rm = TRUE),
                       sd(per_silt, na.rm = TRUE) / sqrt(n)),
    
    pH = sprintf("%.2f ± %.2f",
                 mean(pH, na.rm = TRUE),
                 sd(pH, na.rm = TRUE) / sqrt(n)),
    
    BD_g.cm3 = sprintf("%.2f ± %.2f",
                       mean(BD_g.cm3, na.rm = TRUE),
                       sd(BD_g.cm3, na.rm = TRUE) / sqrt(n)),
    
  biomass_gm2 = sprintf("%.2f ± %.2f",
                       mean(biomass_gm2, na.rm = TRUE),
                       sd(biomass_gm2, na.rm = TRUE) / sqrt(n)),
    
    .groups = "drop"
  )

soc_flex <- flextable(soc_table_fmt) %>%
  set_header_labels(
    PlantingType.y = "Practice",
     climate = "Climate",
    trt = "trt",
    bulkperc1 = "Bulk C (%)",
    corrcpom = "cPOM (%)",
    maomperc = "MAOM (%)",
    soc_Mg.ha = "SOC (Mg ha⁻¹)",
    poc_stock = "POC stock (Mg ha⁻¹)",
    maoc_stock = "MAOC stock (Mg ha⁻¹)",
    fmaoc = "fMAOC",
    per_sand = "Sand (%)",
    per_clay = "Clay (%)",
    per_silt = "Silt (%)",
    pH = "pH",
    BD_g.cm3 = "Bulk density (g cm⁻³)",
     biomass_gm2 = "Herbaceous biomass (g m⁻²)"
  ) %>%
  align(align = "center", part = "all") %>%
  autofit()

save_as_docx(soc_flex, path = "cons_summary_clime_only.docx")


```

Table for canopy
```{r}
## table by canopy 
soc_summary <- fcanopy %>%
  group_by( PlantingType.y,climate, trt) %>%
  summarise(
    n = n(),
    bulkperc_mean = mean(bulkperc, na.rm = TRUE),
    bulkperc_se   = sd(bulkperc, na.rm = TRUE) / sqrt(n),
    
    corrcpom_mean = mean(corrcpom, na.rm = TRUE),
    corrcpom_se   = sd(corrcpom, na.rm = TRUE) / sqrt(n),
    
    maomperc_mean = mean(maomperc, na.rm = TRUE),
    maomperc_se   = sd(maomperc, na.rm = TRUE) / sqrt(n),
    
    .groups = "drop"
  )

soc_summary

write.csv(soc_summary, '/Users/ashban/OneDrive - Colostate/Documents/CSU/Paustian/Point Blue/Land change analysis/working_land/table_climate_only_wsl.csv')

```

Boxplot of all sites (updated 9/12)
```{r}
## boxplot of all sites SOC

ggplot(soc_com, aes(x = trt, y = bulkperc1, fill = trt)) +
  geom_boxplot() +
  facet_wrap(
    ~ pair,
    scales = "free",
    labeller = as_labeller(function(x) paste("Pair", match(x, unique(soc_com$pair))))
  ) +
  theme_bw() +
  xlab(NULL) +
  ylab("SOC (%)") +
  ggtitle("SOC by Treatment for Each Pair") +
  scale_fill_manual(
    values = c("C" = "grey85", "T" = "grey40"),
    labels = c("C" = "Reference", "T" = "Restored")
  ) +
  theme(
    strip.text = element_text(size = 6),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )


```

Graph of climates by soc
```{r}

library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)
install.packages("ggpattern")
library(ggpattern)

# Summarize all three variables by climate
soc_summary <- climate %>%
  mutate(climate = recode(climate,
                          "DRY" = "Dry",
                          "WET" = "Mesic")) %>%
  group_by(climate) %>%
  summarise(
    mean_poc  = mean(poc_stock, na.rm = TRUE),
    se_poc    = sd(poc_stock, na.rm = TRUE) / sqrt(sum(!is.na(poc_stock))),
    mean_maoc = mean(maoc_stock, na.rm = TRUE),
    se_maoc   = sd(maoc_stock, na.rm = TRUE) / sqrt(sum(!is.na(maoc_stock))),
    mean_soc  = mean(soc_Mg.ha, na.rm = TRUE),
    se_soc    = sd(soc_Mg.ha, na.rm = TRUE) / sqrt(sum(!is.na(soc_Mg.ha)))
  )

# Reshape all three to long format
plot_df <- soc_summary %>%
  pivot_longer(
    cols = c(mean_poc, mean_maoc, mean_soc),
    names_to = "fraction",
    values_to = "mean"
  ) %>%
  mutate(
    se = case_when(
      fraction == "mean_poc"  ~ se_poc,
      fraction == "mean_maoc" ~ se_maoc,
      fraction == "mean_soc"  ~ se_soc
    ),
    fraction = recode(fraction,
                      "mean_poc"  = "POC",
                      "mean_maoc" = "MAOC",
                      "mean_soc"  = "SOC"),
    # Control bar order within each climate group
    fraction = factor(fraction, levels = c("POC", "MAOC", "SOC"))
  ) %>%
  select(climate, fraction, mean, se)

# Plot
ggplot(plot_df, aes(x = climate, y = mean, fill = fraction, pattern = fraction)) +
  ggpattern::geom_bar_pattern(
    stat          = "identity",
    position      = position_dodge(width = 0.75),
    color         = "black",
    width         = 0.65,
    pattern_fill  = "black",
    pattern_color = "black",
    pattern_angle = 45,
    pattern_density = 0.3,
    pattern_spacing = 0.03
  ) +
  geom_errorbar(
    aes(ymin = mean - se, ymax = mean + se),
    position = position_dodge(width = 0.75),
    width    = 0.2,
    linewidth = 0.7
  ) +
  scale_fill_manual(
    values = c("POC"  = "white",
               "MAOC" = "white",
               "SOC"  = "gray50")  # SOC solid gray to distinguish it
  ) +
  ggpattern::scale_pattern_manual(
    values = c("POC"  = "none",     # solid white
               "MAOC" = "stripe",   # hatched
               "SOC"  = "none")     # solid gray
  ) +
  labs(
    x        = "Climate",
    y        = expression("Stock (Mg ha"^-1*")"),
    fill     = NULL,
    pattern  = NULL,
    title    = "SOC Fractions by Climate"
  ) +
  theme_bw() +
  theme(
    panel.grid.major  = element_blank(),
    panel.grid.minor  = element_blank(),
    axis.text         = element_text(size = 12),
    axis.title        = element_text(size = 14),
    legend.title      = element_text(size = 11),
    legend.text       = element_text(size = 10),
    plot.margin       = margin(t = 30,  # top
                               r = 30,  # right
                               b = 30,  # bottom
                               l = 30),  # left, in points
    legend.position   = "top"
    
  ) +
  guides(fill    = guide_legend(override.aes = list(pattern = c("none", "stripe", "none"))),
         pattern = "none")  # suppress duplicate legend

```
Graph of canopy 
```{r}


# Install if needed
install.packages("ggsignif")
library(ggsignif)

# Summarize by canopy
canopy_summary <- fcanopynona %>%
  group_by(Canopy) %>%
  summarise(
    mean_soc = mean(bulkperc1, na.rm = TRUE),
    se_soc   = sd(bulkperc1, na.rm = TRUE) / sqrt(sum(!is.na(bulkperc1))),
    .groups  = "drop"
  )

# Plot
plot<- ggplot(canopy_summary, aes(x = Canopy, y = mean_soc, fill = Canopy, pattern = Canopy)) +
  ggpattern::geom_bar_pattern(
    stat            = "identity",
    color           = "black",
    width           = 0.6,
    pattern_fill    = "gray40",
    pattern_color   = "gray40",
    pattern_angle   = 45,
    pattern_density = 0.3,
    pattern_spacing = 0.03
  ) +
  geom_errorbar(
    aes(ymin = mean_soc - se_soc,
        ymax = mean_soc + se_soc),
    width     = 0.1,
    linewidth = 0.7
  ) +
  geom_signif(
    comparisons = list(c("Outside Canopy", "Under Canopy")),
    annotations = "**",
    tip_length  = 0.02,
    textsize    = 6,
    vjust       = 0.5,
    y_position  = 2.2    # adjust this number up/down as needed
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +  # ← added here
  scale_fill_manual(
    values = c("Outside Canopy" = "gray85",
               "Under Canopy"   = "gray45")
  ) +
  ggpattern::scale_pattern_manual(
    values = c("Outside Canopy" = "stripe",
               "Under Canopy"   = "none")
  ) +
  labs(
    x = NULL,
    y = "SOC (%)"
  ) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text        = element_text(size = 12),
    axis.title       = element_text(size = 14),
    legend.position  = "none",
    plot.margin      = margin(t = 20, r = 20, b = 20, l = 20)
  )

plot
```
Graph of carbon over time by practice 
```{r fig.width=12, fig.height=6}


# 1. Clean and recode the main dataframe
soc <- socfnona_0 %>%
  mutate(climate = case_match(climate,
                              "DRY" ~ "Dry",
                              "WET" ~ "Mesic",
                              .default = climate))

#### Plot ####
# 2. Build the plotting dataframe using the 'soc' object we just created
plot_df <- soc %>%
  select(PlantingType.y, climate, age_2023, soc_Mg.ha, maoc_stock, poc_stock) %>%
  pivot_longer(
    cols      = c(soc_Mg.ha, maoc_stock, poc_stock),
    names_to  = "fraction",
    values_to = "stock"
  ) %>%
  mutate(
    fraction = case_match(fraction,
      "soc_Mg.ha"  ~ "SOC Stock\n(Mg C ha-1 Yr-1)",
      "maoc_stock" ~ "MAOC Stock\n(Mg C ha-1 Yr-1)",
      "poc_stock"  ~ "POC Stock\n(Mg C ha-1 Yr-1)",
      .default     = fraction
    ),
    fraction = factor(fraction, levels = c(
      "SOC Stock\n(Mg C ha-1 Yr-1)",
      "MAOC Stock\n(Mg C ha-1 Yr-1)",
      "POC Stock\n(Mg C ha-1 Yr-1)"
    )),
    climate = case_match(climate,
      "Dry"   ~ "Dry Regions",
      "Mesic" ~ "Mesic Regions",
      .default = climate
    )
  )

    

options(repr.plot.width = 10, repr.plot.height = 6)
# Plot
ggplot(plot_df, aes(x = age_2023, y = stock, 
                    color = PlantingType.y, 
                    shape = PlantingType.y)) +
  geom_point(alpha = 0.4, size = 2) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", linewidth = 0.8) +
  facet_grid(fraction ~ climate, scales = "free_y") +
  scale_color_manual(values = c(
    "Riparian" = "maroon",   # replace Type1/2/3 with your actual PlantingType labels
    "UplandOak" = "#56B4E9",
    "RangePlanting" = "seagreen3"
  )) +
  labs(
    x     = "Years of practice",
    y     = NULL,
    color = NULL,
    shape = NULL
  ) +
  theme_bw() +
  theme(
    panel.grid.minor  = element_blank(),
    strip.background  = element_blank(),
    strip.text        = element_text(size = 11),
    axis.text         = element_text(size = 10),
    axis.title        = element_text(size = 12),
    legend.position   = "bottom"
  )

```
Table soc for climate (mesic/dry)
````{r}
library(dplyr)
library(writexl)

# Means and SE by climate for each fraction
climate_summary <- socfnona %>%
  group_by(climate) %>%
  summarise(
    # SOC
    SOC_mean = round(mean(soc_Mg.ha, na.rm = TRUE), 2),
    SOC_se   = round(sd(soc_Mg.ha, na.rm = TRUE) / sqrt(sum(!is.na(soc_Mg.ha))), 2),
    # MAOC
    MAOC_mean = round(mean(maoc_stock, na.rm = TRUE), 2),
    MAOC_se   = round(sd(maoc_stock, na.rm = TRUE) / sqrt(sum(!is.na(maoc_stock))), 2),
    # POC
    POC_mean = round(mean(poc_stock, na.rm = TRUE), 2),
    POC_se   = round(sd(poc_stock, na.rm = TRUE) / sqrt(sum(!is.na(poc_stock))), 2),
    n = n()
  )

# F-statistics from climate term in each model
soc_anova  <- anova(aov_soc)
maoc_anova <- anova(aov_maoc)
poc_anova  <- anova(aov_poc)

# Extract climate F and P values
f_table <- data.frame(
  Fraction = c("SOC", "MAOC", "POC"),
  F_stat   = round(c(
    soc_anova["climate", "F value"],
    maoc_anova["climate", "F value"],
    poc_anova["climate", "F value"]), 2),
  P_value  = round(c(
    soc_anova["climate", "Pr(>F)"],
    maoc_anova["climate", "Pr(>F)"],
    poc_anova["climate", "Pr(>F)"]), 4)
)

# Print to check
print(climate_summary)
print(f_table)

# Export
write_xlsx(list(
  Climate_Summary = as.data.frame(climate_summary),
  F_statistics    = f_table),
  "climate_carbon_table.xlsx")



````


```{r}
# Step 1 - collapse pseudoreplicates to site level
site_means <- socfnona %>%
  group_by(ranch.pract) %>%
  summarise(
    SOC  = mean(soc_Mg.ha, na.rm = TRUE),
    MAOC = mean(maoc_stock, na.rm = TRUE),
    POC  = mean(poc_stock, na.rm = TRUE),
    .groups = "drop"
  )

# Confirm correct number of sites
nrow(site_means)

# Step 2 - variability across site means
data_var <- data.frame(
  Fraction = c("SOC", "MAOC", "POC"),
  Mean     = c(mean(site_means$SOC),  mean(site_means$MAOC),  mean(site_means$POC)),
  SD       = c(sd(site_means$SOC),    sd(site_means$MAOC),    sd(site_means$POC)),
  Min      = c(min(site_means$SOC),   min(site_means$MAOC),   min(site_means$POC)),
  Max      = c(max(site_means$SOC),   max(site_means$MAOC),   max(site_means$POC)),
  Fold     = c(max(site_means$SOC) / min(site_means$SOC),
               max(site_means$MAOC) / min(site_means$MAOC),
               max(site_means$POC) / min(site_means$POC)),
  CV       = c(sd(site_means$SOC)  / mean(site_means$SOC)  * 100,
               sd(site_means$MAOC) / mean(site_means$MAOC) * 100,
               sd(site_means$POC)  / mean(site_means$POC)  * 100)
) %>%
  mutate(across(where(is.numeric), ~ round(., 2)))

write_xlsx(data_var, "carbon_variability_summary.xlsx")
```

graph of conservation practice distribution by climate
```{r}
#conservation practice distribution by climate 
head(socfnona)
# Clean up labels if needed
fcanopy <- socfnona %>%
  mutate(
    climate = recode(climate, "DRY" = "Dry", "WET" = "Mesic"),
    climate = factor(climate, levels = c("Dry", "Mesic")),
    climate = recode(climate, "T" = "Restored", "C" = "Reference"),
  )

ggplot(fcanopy, aes(x = climate, y = bulkperc1, fill = trt)) +  # replace trt with your reference/restored column name
  geom_boxplot(
    outlier.shape  = 16,
    outlier.size   = 1.5,
    outlier.colour = "gray40",
    width          = 0.6,
    position       = position_dodge(width = 0.75)
  ) +
  facet_wrap(~ PlantingType.y, nrow = 1) +  # one panel per practice
  scale_fill_manual(
    values = c("C" = "gray85",   # light gray = reference
               "T"  = "gray40"),  # dark gray = restored
    name   = NULL
  ) +
  scale_x_discrete(labels = c("Dry", "Mesic")) +
  labs(
    x = NULL,
    y = "SOC (%)"
  ) +
  theme_bw() +
  theme(
    panel.grid.major  = element_blank(),
    panel.grid.minor  = element_blank(),
    strip.background  = element_rect(fill = "gray90", color = "black"),
    strip.text        = element_text(size = 12, face = "bold"),
    axis.text         = element_text(size = 11),
    axis.title        = element_text(size = 13),
    legend.position   = "right",
    legend.text       = element_text(size = 11),
    plot.margin       = margin(t = 20, r = 20, b = 20, l = 20)
  )




```

# SOC ~ trt*year + (1|ranch/pair/plot) - graphs for practices 
```{r}

# ── Models: all practices together with climate and PlantingType.y ────────────
soc_time  <- lmer(log(soc_Mg.ha)  ~ trt * age_2023 * climate * PlantingType.y + per_clay + 
                    (1 | RanchName.x/pair/ranch.pract), data = socfnona)
maoc_time <- lmer(log(maoc_stock) ~ trt * age_2023 * climate * PlantingType.y + per_clay + 
                    (1 | RanchName.x/pair/ranch.pract), data = socfnona)
poc_time  <- lmer(log(poc_stock)  ~ trt * age_2023 * climate * PlantingType.y + per_clay + 
                    (1 | RanchName.x/pair/ranch.pract), data = socfnona)

# ── Check interaction significance ───────────────────────────────────────────
anova(soc_time)
anova(maoc_time)
anova(poc_time)

# ── Extract slopes by trt, climate, and PlantingType.y ───────────────────────
get_trends <- function(model, fraction) {
  ci <- emtrends(model, ~ trt | climate + PlantingType.y, var = "age_2023") %>% 
    as.data.frame()
  pv <- emtrends(model, ~ trt | climate + PlantingType.y, var = "age_2023") %>% 
    test() %>% 
    as.data.frame() %>%
    select(trt, climate, PlantingType.y, t.ratio, p.value)
  left_join(ci, pv, by = c("trt", "climate", "PlantingType.y")) %>%
    mutate(Fraction = fraction)
}

all_trends <- bind_rows(
  get_trends(soc_time,  "SOC"),
  get_trends(maoc_time, "MAOC"),
  get_trends(poc_time,  "POC")
) %>%
  select(Fraction, climate, PlantingType.y, trt, age_2023.trend, SE, 
         lower.CL, upper.CL, t.ratio, p.value) %>%
  rename(
    Climate   = climate,
    Practice  = PlantingType.y,
    Treatment = trt,
    Slope     = age_2023.trend,
    P.value   = p.value
  ) %>%
  mutate(across(where(is.numeric), ~ round(.x, 4)))

# ── Model fit statistics ──────────────────────────────────────────────────────
model_fit <- data.frame(
  Response = c("SOC", "MAOC", "POC"),
  AIC      = c(AIC(soc_time), AIC(maoc_time), AIC(poc_time)),
  BIC      = c(BIC(soc_time), BIC(maoc_time), BIC(poc_time))
) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

# ── Plot ──────────────────────────────────────────────────────────────────────
plot_df <- socfnona %>%
  mutate(
    climate = case_match(climate,
      "DRY" ~ "Dry Regions",
      "WET" ~ "Mesic Regions",
      .default = climate
    )
  ) %>%
  select(trt, PlantingType.y, climate, age_2023, soc_Mg.ha, maoc_stock, poc_stock) %>%
  pivot_longer(
    cols      = c(soc_Mg.ha, maoc_stock, poc_stock),
    names_to  = "fraction",
    values_to = "stock"
  ) %>%
  mutate(
    fraction = case_match(fraction,
      "soc_Mg.ha"  ~ "SOC Stock\n(Mg C ha⁻¹ Yr⁻¹)",
      "maoc_stock" ~ "MAOC Stock\n(Mg C ha⁻¹ Yr⁻¹)",
      "poc_stock"  ~ "POC Stock\n(Mg C ha⁻¹ Yr⁻¹)"
    ),
    fraction = factor(fraction, levels = c(
      "SOC Stock\n(Mg C ha⁻¹ Yr⁻¹)",
      "MAOC Stock\n(Mg C ha⁻¹ Yr⁻¹)",
      "POC Stock\n(Mg C ha⁻¹ Yr⁻¹)"
    ))
  )


practice_colors <- c(
  "Riparian"      = "maroon",
  "UplandOak"     = "#56B4E9",
  "RangePlanting" = "seagreen3"
)

for (practice in unique(plot_df$PlantingType.y)) {
  
  p <- ggplot(
    subset(plot_df, PlantingType.y == practice),
    aes(
      x = age_2023,
      y = stock,
      linetype = trt,
      shape = trt
    )
  ) +
    geom_point(
      color = practice_colors[practice],
      alpha = 0.4,
      size = 2
    ) +
    geom_smooth(
      color = practice_colors[practice],
      method = "lm",
      se = FALSE,
      linewidth = 0.8
    ) +
    facet_grid(
      fraction ~ climate,
      scales = "free_y"
    ) +
    scale_linetype_manual(
      values = c(
        "T" = "solid",
        "C" = "dashed"
      ),
      labels = c(
        "T" = "Treatment",
        "C" = "Control"
      )
    ) +
    scale_shape_manual(
      values = c(
        "T" = 16,
        "C" = 17
      ),
      labels = c(
        "T" = "Treatment",
        "C" = "Control"
      )
    ) +
    labs(
      title = practice,
      x = "Years of practice",
      y = NULL,
      linetype = NULL,
      shape = NULL
    ) +
    theme_bw() +
    theme(
      panel.grid.minor = element_blank(),
      strip.background = element_blank(),
      strip.text = element_text(size = 11),
      plot.title = element_text(size = 14, face = "bold"),
      axis.text = element_text(size = 10),
      axis.title = element_text(size = 12),
      legend.position = "bottom"
    )
  
  print(p)
}

get_slope_comparisons <- function(model, fraction) {

  trends <- emtrends(
    model,
    ~ trt | climate + PlantingType.y,
    var = "age_2023"
  )

  contrast(
    trends,
    method = list("T - C" = c(-1, 1))
  ) %>%
    as.data.frame() %>%
    mutate(Fraction = fraction)

}

slope_comparisons <- bind_rows(
  get_slope_comparisons(soc_time, "SOC"),
  get_slope_comparisons(maoc_time, "MAOC"),
  get_slope_comparisons(poc_time, "POC")
)

slope_comparisons

write.csv(
  slope_comparisons,
  "slope_comparisons.csv",
  row.names = FALSE
)


```


# ── Calculate T - C difference within each pair ───────────────────────────────
```{r}

diff_df <- socfnona %>%
  select(pair, RanchName.x, PlantingType.y, climate, trt, age_2023,
         soc_Mg.ha, maoc_stock, poc_stock) %>%
  pivot_longer(
    cols      = c(soc_Mg.ha, maoc_stock, poc_stock),
    names_to  = "fraction",
    values_to = "stock"
  ) %>%
  group_by(pair, PlantingType.y, climate, fraction, age_2023) %>%
  summarise(
    diff = stock[trt == "T"] - stock[trt == "C"],
    .groups = "drop"
  ) %>%
  mutate(
    fraction = case_match(fraction,
      "soc_Mg.ha"  ~ "SOC Stock\n(Mg C ha⁻¹)",
      "maoc_stock" ~ "MAOC Stock\n(Mg C ha⁻¹)",
      "poc_stock"  ~ "POC Stock\n(Mg C ha⁻¹)"
    ),
    fraction = factor(fraction, levels = c(
      "SOC Stock\n(Mg C ha⁻¹)",
      "MAOC Stock\n(Mg C ha⁻¹)",
      "POC Stock\n(Mg C ha⁻¹)"
    )),
    climate = case_match(climate,
      "DRY" ~ "Dry Regions",
      "WET" ~ "Mesic Regions",
      .default = climate
    )
  )

# ── Plot the difference over time ─────────────────────────────────────────────
ggplot(diff_df, aes(x = age_2023, y = diff, color = PlantingType.y)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.6) +
  geom_point(alpha = 0.4, size = 2) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  facet_grid(fraction ~ climate, scales = "free_y") +
  scale_color_manual(values = c(
    "Riparian"      = "maroon",
    "UplandOak"     = "#56B4E9",
    "RangePlanting" = "seagreen3"
  )) +
  labs(
    x     = "Years of practice",
    y     = "Treatment − Control (Mg C ha⁻¹)",
    color = NULL
  ) +
  theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text       = element_text(size = 11),
    axis.text        = element_text(size = 10),
    axis.title       = element_text(size = 12),
    legend.position  = "bottom"
  )


```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
